---
title: "ebird-data-open"
output: html_document
---

family of interest: Hirundinidae

Progne subis (PUMA)
Tachycineta bicolor (TRES)
Stelgidopteryx serripennis (NRWS)
Riparia riparia (BANS)
Petrochelidon pyrrhonota (CLSW)
Hirundo rustica (BARS)

```{r}
install.packages("data.table")
install.packages("tidyverse")
library(tidyverse)
library(data.table)
```

#Reading the checklist file
```{r}
ebird_2003 <- fread("C:/Temp/ERD2002SS/2003/checklists.csv", select = c("SAMPLING_EVENT_ID", "LOC_ID", "LATITUDE", "LONGITUDE", "YEAR", "MONTH", "DAY", "TIME",
                                                                   "COUNTRY", "STATE_PROVINCE", "COUNTY", "COUNT_TYPE", "EFFORT_HRS", "EFFORT_DISTANCE_KM", 
                                                                   "EFFORT_AREA_HA", "OBSERVER_ID", "NUMBER_OBSERVERS", "GROUP_ID", "PRIMARY_CHECKLIST_FLAG", 
                                                                   "Progne_subis", "Tachycineta_bicolor", "Stelgidopteryx_serripennis", "Riparia_riparia", 
                                                                   "Petrochelidon_pyrrhonota", "Hirundo_rustica"))
```

To confirm it's the correct number of rows, check the number of rows in the csv file (without opening it). Open the command line prompt and type 
`find /c /v "A String that is extremely unlikely to occur" filename`. Checks out!

#Subsetting
```{r}
ebird_2003 <- ebird_2003[!(ebird_2003$Progne_subis == 0 & ebird_2003$Tachycineta_bicolor == 0 & ebird_2003$Stelgidopteryx_serripennis == 0 & 
                             ebird_2003$Riparia_riparia == 0 & ebird_2003$Petrochelidon_pyrrhonota == 0 & ebird_2003$Hirundo_rustica == 0)]
```

#Write into new csv
```{r}
fwrite(ebird_2003, "ebird_2003_small.csv")
```

#Try for the rest of the checklist files

```{r}
ebird_2002 <- fread("C:/Temp/ERD2016SS/2002/checklists.csv", select = c("SAMPLING_EVENT_ID", "LOC_ID", "LATITUDE", "LONGITUDE", "YEAR", "MONTH", "DAY", "TIME",
                                                                   "COUNTRY", "STATE_PROVINCE", "COUNTY", "COUNT_TYPE", "EFFORT_HRS", "EFFORT_DISTANCE_KM", 
                                                                   "EFFORT_AREA_HA", "OBSERVER_ID", "NUMBER_OBSERVERS", "GROUP_ID", "PRIMARY_CHECKLIST_FLAG", 
                                                                   "Progne_subis", "Tachycineta_bicolor", "Stelgidopteryx_serripennis", "Riparia_riparia", 
                                                                   "Petrochelidon_pyrrhonota", "Hirundo_rustica"))
ebird_2002 <- ebird_2002[!(ebird_2002$Progne_subis == 0 & ebird_2002$Tachycineta_bicolor == 0 & ebird_2002$Stelgidopteryx_serripennis == 0 & 
                             ebird_2002$Riparia_riparia == 0 & ebird_2002$Petrochelidon_pyrrhonota == 0 & ebird_2002$Hirundo_rustica == 0)]
fwrite(ebird_2002, "ebird_2002_small.csv")
```

fread() for 2000-six + 2000-nine onwards had a weird output. did something go wrong? these files all had many rows (>100 000). files seem ok though

|--------------------------------------------------|
|==================================================|


