---
title: "exploratory analysis"
output:
  pdf_document: default
  html_document: default
---

# Let's try the breeding bird dataset

```{r message=FALSE, warning=FALSE}
library(tidyverse)
birds <- read.csv("Ontario.csv")
checklist <- read.csv("checklist.csv")
```

##See how total abundance changes year by year
```{r}
birds$family <- checklist$Family[match(birds$AOU, checklist$AOU)]
birds$genus <- checklist$Genus[match(birds$AOU, checklist$AOU)]
birds$species <- checklist$common_name[match(birds$AOU, checklist$AOU)]
birds$seq <- checklist$Seq[match(birds$AOU, checklist$AOU)]

yearly <- birds %>%
  group_by(seq, family, genus, species, Year) %>% 
  summarize(annual_spp_abundance = mean(SpeciesTotal)) %>% 
  arrange(seq)
```

##Plot!

Waterfowl population
```{r}
yearly %>% 
  filter(family == "Anatidae") %>% 
  ggplot(aes(x = Year, y = annual_spp_abundance, colour = genus)) + geom_point() +
  geom_smooth()
#huge increase in geese!
```

Just out of curiosity, let's look at individual species (and some species at risk)

```{r}
yearly %>% 
  filter(family == "Hirundinidae") %>% 
  ggplot(aes(x = Year, y = annual_spp_abundance, colour = species)) + geom_point() +
  geom_smooth()
```

```{r}
yearly %>% 
  filter(species == "Bobolink") %>% 
  ggplot(aes(x = Year, y = annual_spp_abundance)) + geom_point() +
  geom_smooth()
```

##Questions

It would be interesting to look at regional differences in population dynamics. 
Are bird populations increasing/decreasing across their full range or regionally? Any range shifts?

There are also files with environmental and weather conditions. Can noise/level of human activity/weather explain these perceived changes in abundance?

Use multiple datasets to study dynamics of a focal species/group?

# Let's try another dataset: Marine predator and prey body sizes

```{r}
marine <- read.csv("Marine.csv")
glimpse(marine)
str(marine)
```

Rich dataset! Can ask lots of questions. But this will need lots of cleaning. 

```{r}
marine %>% 
  ggplot(aes(x = Prey.mass), colour = Predator) + geom_histogram()
```

















































