---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)

birds <- read_csv("Swallows Data Expanded.csv")
#View(birds)
```

#Q Plots for Swallow Total vs Temps 
```{r}
qplot(StartTemp, SpeciesTotal, data = birds)
```

```{r}
qplot(EndTemp, SpeciesTotal, data = birds)
```

#Q Plots for Bird Species Total vs Temps
```{r}
qplot(StartTemp, TotalSpp, data = birds)
```

```{r}
qplot(EndTemp, TotalSpp, data = birds)
```

#Compiled ggplots of the above qplots
```{r}
ggplot(birds, aes(y = SpeciesTotal))+
  geom_point(aes(x = StartTemp), colour = "orange")+
  geom_point(aes(x = EndTemp), colour = "blue", alpha = 0.25)+
  xlab("Temp")
```

```{r}
ggplot(birds, aes(y = TotalSpp))+
  geom_point(aes(x = StartTemp), colour = "purple")+
  geom_point(aes(x = EndTemp), colour = "pink", alpha = 0.25)+
  xlab("Temp")
```

#Plot for each Swallow species based on Latitude/Longitude
This graph has potential for making a heat map! Since you can basically almost see the outline of Southern Ontario!
```{r}
ggplot(birds, aes(x = Longitude, y = Latitude))+
  geom_point(aes(colour = species))+
  facet_wrap(~species)+
  theme(axis.text.x = element_text(size = 9, angle = 45, hjust = 1))
```

#Plot for Swallows vs Sky Conditions
```{r}
ggplot(birds, aes(StartSky))+
  geom_point(aes(y = SpeciesTotal), colour = "blue4")+
  facet_wrap(~species)
```

```{r}
ggplot(birds, aes(EndSky))+
  geom_point(aes(y = SpeciesTotal), colour = "cornflowerblue")+
  facet_wrap(~species)
```

```{r}
ggplot(birds, aes(y = SpeciesTotal))+
  geom_point(aes(x = StartSky), colour = "blue4")+
  geom_point(aes(x = EndSky), colour = "cornflowerblue", alpha = 0.25)+
  xlab("Sky Conditions")+
  facet_wrap(~species)
```


#Plot for All birds vs Sky Conditions
```{r}
ggplot(birds, aes(StartSky))+
  geom_point(aes(y = TotalSpp), colour = "blue4")+
  facet_wrap(~species)
```

```{r}
ggplot(birds, aes(EndSky))+
  geom_point(aes(y = TotalSpp), colour = "cornflowerblue")+
  facet_wrap(~species)
```

```{r}
ggplot(birds, aes(y = TotalSpp))+
  geom_point(aes(x = StartSky), colour = "blue4")+
  geom_point(aes(x = EndSky), colour = "cornflowerblue", alpha = 0.25)+
  xlab("Sky Conditions")+
  facet_wrap(~species)
```

#Plot for Swallows vs Wind Conditions
```{r}
ggplot(birds, aes(y = SpeciesTotal))+
  geom_point(aes(x = StartWind), colour = "greenyellow")+
  geom_point(aes(x = EndWind), colour = "honeydew4", alpha = 0.25)+
  xlab("Wind Conditions")+
  facet_wrap(~species)
```

#Just some messing around with some graphs
```{r}
birds$Car_Total <- as.numeric(birds$Car_Total)

qplot(SpeciesTotal, Car_Total, data = birds)+
  geom_point()
```

```{r}
ggplot(birds, aes(Car_Total))+
  geom_point(aes(y = TotalSpp), colour = "hotpink")+
  geom_point(aes(y = SpeciesTotal), colour = "hotpink4", alpha = 0.25)+
  facet_wrap(~species)
```







