---
title: "R Code (Richard)"
output: html_notebook
---

# Create a mega-dataset for all the potential data we need (Nov.15-20)

```{r}
library(tidyverse)
checklist <- read_csv("checklist.csv")
vehicle <- read_csv("VehicleData.csv")
weather <- read_csv("weather.csv")
routes <- read_csv("routes.csv")
```

## Clean up vehicle data

```{r}
# Create new columns that contain total number of cars per 10 routes -- to be matching with Ontario.csv

vehicle1 <- vehicle %>%
  filter(CountryNum == 124, StateNum == 68) # Looking at data in Ontario

# Calculate sum of cars by every 10 stops (could not mutate because there are 10 columns)
for (row in c(1:nrow(vehicle1))){
    vehicle1[row, 108] <- sum(vehicle1[row,8:17])
    vehicle1[row, 109] <- sum(vehicle1[row,18:27])
    vehicle1[row, 110] <- sum(vehicle1[row,28:37])
    vehicle1[row, 111] <- sum(vehicle1[row,38:47])
    vehicle1[row, 112] <- sum(vehicle1[row,48:57])
}

# Change column names of the new columns
colnames(vehicle1)[colnames(vehicle1)=="V108"] <- "Car1_10"
colnames(vehicle1)[colnames(vehicle1)=="V109"] <- "Car11_20"
colnames(vehicle1)[colnames(vehicle1)=="V110"] <- "Car21_30"
colnames(vehicle1)[colnames(vehicle1)=="V111"] <- "Car31_40"
colnames(vehicle1)[colnames(vehicle1)=="V112"] <- "Car41_50"

# Mutate a new column for the sum of cars from 1-50
vehicle1 <- vehicle1 %>% 
  mutate(Car_Total = Car1_10 + Car11_20 + Car21_30 + Car31_40 + Car41_50)

# Mutate a new column for matching purposes
vehicle1 <- mutate(vehicle1, RouteYear = as.numeric(Route) * 10000 + as.numeric(Year))

# Reorder the new columns to front of the dataframe for easy visualization
vehicle2 <- vehicle1[, c(1:6, 114, 7, 113, 108:112, 8:107)]
```

## Clean up bird abundance data (using checklist.csv)

```{r}
birds <- read_csv("Ontario.csv")
birds$AOU <- as.integer(birds$AOU)

birds$family <- checklist$Family[match(birds$AOU, checklist$AOU)]
birds$genus <- checklist$Genus[match(birds$AOU, checklist$AOU)]
birds$species <- checklist$common_name[match(birds$AOU, checklist$AOU)]
birds$seq <- checklist$Seq[match(birds$AOU, checklist$AOU)]

# Mutate a new column for matching purposes
birds <- mutate(birds, RouteYear = as.numeric(Route) * 10000 + as.numeric(Year))

# Reorder for easy visualization
birds1 <- birds[, c(19, 17, 13:14, 8:12, 1:7, 15:16, 18)]
```

## Clean up weather data

```{r}
weather1 <- weather %>%
  filter(CountryNum == 124, StateNum == 68) # Looking at data in Ontario

# Mutate a new column for matching purposes
weather1 <- mutate(weather1, RouteYear = as.numeric(Route) * 10000 + as.numeric(Year))
```

## Clean up route data

```{r}
routes1 <- routes %>%
  filter(CountryNum == 124, StateNum == 68) # Looking at data in Ontario
```

## Add count of routes in each year

```{r}
# These are just Caroline's codes, some figure
birds$Route <- as.factor(birds$Route)
table1 <- birds1 %>% 
  group_by(Year, Route) %>% 
  count(Year, Route) %>% 
  select(Year, Route)
table1 <- table1 %>% 
  group_by(Year) %>% 
  tally()
ggplot(table1, aes(x = Year, y = n)) + geom_point()
```

## Combine 4+1 datasets: bird abundance (including checklist), vehicle data, weather, and route details + table 1 (Caroline's code)

```{r}
expanded_df <- merge(birds1, vehicle2, by = c("RouteYear", "RouteDataID", "CountryNum", "StateNum", "Route", "RPID", "Year"), all = TRUE)
expanded_df <- merge(expanded_df, weather1, by = c("RouteYear", "RouteDataID", "CountryNum", "StateNum", "Route", "RPID", "Year"), all = TRUE)
expanded_df <- merge(expanded_df, routes1, by = c("CountryNum", "StateNum", "Route"), all = TRUE)
expanded_df <- merge(expanded_df, table1, by = "Year")

expanded_df <- expanded_df %>% 
  arrange(Year)

head(expanded_df)

# Reorder columns to make life easier
expanded_df1 <- expanded_df[, c(5, 8, 9:10, 20:21, 127:151, 11:15, 22:126, 1:4, 6:7, 16:19)]

# Rename "n"
colnames(expanded_df1)[colnames(expanded_df1)=="n"] <- "NoOfRouteInYear"
head(expanded_df1)
```

## Output expanded .csv files

```{r}
write_csv(expanded_df1, "Ontario Bird Data Expanded.csv")
```

```{r}
expanded_df2 <- expanded_df1 %>% 
  filter(family == "Hirundinidae")

write_csv(expanded_df2, "Swallows only - Ontario Bird Data Expanded.csv")
```


# Cleanup Mega-dataset (Nov.20-22)

```{r}
library(tidyverse)

swallows <- read_csv("Swallows only - Ontario Bird Data Expanded.csv", col_types = cols(TempScale = "c"))
```

## Convert all Fahrenheit to Celcius

```{r}
swallows$StartTemp <- as.numeric(swallows$StartTemp)
swallows$EndTemp <- as.numeric(swallows$EndTemp)

# str(swallows$TempScale)
unique(swallows$TempScale)

# Actual TempScale containts: "F", "C", "NULL" and "NA"
# "NA"s do not have any temperature recordings (displayed as 0, I will change them to NA)
# There are two cases of "NULL"s, one has temperatures in Fahrenheit [will change to Celcius] 
# and the other has "NULL" [will change to NA]

swallows_modified <- swallows %>% 
  # Convert F to C
  mutate(StartT = ifelse(TempScale == "F", (StartTemp - 32) / 1.8, StartTemp)) %>%
  mutate(EndT = ifelse(TempScale == "F", (EndTemp - 32) / 1.8, EndTemp)) %>% 
  
  # Convert NA temps to "NA"
  mutate(StartT1 = ifelse(TempScale == "NA", NA, StartT)) %>%
  mutate(EndT1 = ifelse(TempScale == "NA", NA, EndT)) %>%

  # Convert NULL temps to F or NA
  mutate(StartT2 = ifelse(TempScale == "NULL" & !is.na(StartTemp), (StartTemp - 32) / 1.8, StartT1)) %>%
  mutate(EndT2 = ifelse(TempScale == "NULL" & !is.na(EndTemp), (EndTemp - 32) / 1.8, EndT1)) %>% 
  mutate(StartT3 = ifelse(TempScale == "NULL" & is.na(StartTemp), NA, StartT2)) %>%
  mutate(EndT3 = ifelse(TempScale == "NULL" & is.na(EndTemp), NA, EndT2)) %>% 

  # There are some cases with errorneous recordings, or only have one of the 2 temps, will replace both temps as NA
  mutate(StartT4 = ifelse(abs(StartT3 - EndT3) > 50, NA, StartT3)) %>%
  mutate(EndT4 = ifelse(abs(StartT3 - EndT3) > 50, NA, EndT3))

# swallows_modified <- arrange(swallows_modified, desc(TempScale))
write_csv(swallows_modified, "temp.csv")

```

## Delete confounding columns and rename new start and end temps

```{r}
swallows_modified1 <- swallows_modified[,c(1:10, 13, 160:161, 14:151)]

colnames(swallows_modified1)[colnames(swallows_modified1) == "StartT4"] <- "StartTemp"
colnames(swallows_modified1)[colnames(swallows_modified1) == "EndT4"] <- "EndTemp"

# Remove TempScale column
swallows_modified1 <- swallows_modified1 %>% 
  select(-TempScale)
```

## Select columns with RouteTypeDetailID as 1 (most complete routes) or NA as most are NA

```{r}
swallows_modified2 <- swallows_modified1 %>% 
  filter(RouteTypeDetailID == 1 | is.na(RouteTypeDetailID))

# RouteTypeID and RouteTypeDetailID are now either 1 or NA, no longer relevant
# So they will be moved to the back of dataset
swallows_modified2 <- swallows_modified2[,c(1:27, 30:150, 28:29)]
```

## Output expanded .csv file

```{r}
write_csv(swallows_modified2, "Swallows Data Expanded.csv")
```



# Generate Distibution Plots (Nov.22-29)

```{r}
library(tidyverse)

swallows <- read_csv("Swallows Data Expanded.csv")

swallows <- swallows[,c(139, 1:138, 140:150)]

# Out of 9703 rows, 9608 have SpeciesTotal <= 100, so adjust >100 to 100 for better heatmap
swallows1 <- swallows %>% 
  mutate(SpeciesTotal1 = ifelse(SpeciesTotal > 100, 100, SpeciesTotal))

swallows1 <- swallows1[, c(1, 151, 2:150)]
```

## Draw map of Ontario (Took forever to figure out)

Region of observations: Longitude between -95 and -74, Latitude between 42 and 53

```{r}
# install.packages("maps")
# install.packages("mapdata")
library(maps)
library(mapdata)

# Map of the part of Ontario we are interested in
ontario_map1 <- map_data("world", regions = "Canada") %>% 
  filter(long > -95 & long < -74) %>% 
  filter(lat > 42 & lat < 60)

area_map <- ggplot() + 
  geom_polygon(data = ontario_map1, aes(x = long, y = lat, group = group),
               fill = "#F5F5F5", color = "black") + 
  coord_fixed(1.3) 

# Some major Ontario cities
ont_cities <- canada.cities %>% 
  filter(long > -95 & long < -74) %>% 
  filter(lat > 42 & lat < 60) %>% 
  filter(pop > 110000)

ont_cities %>% arrange(long)
```

## Make distribution plot -- distribution to year (Took a lot of time)

Note: More routes have been observed between 1994-2015 (i.e. >=75).

```{r}
Swallows_temp <- swallows1 %>%
  filter(Year == 1994, species == "Tree Swallow")

# Generate plot title
Plot_Title <- paste("Ontario", unique(Swallows_temp$species), "Distribution:", unique(Swallows_temp$Year) )

# Generate plot 
area_map +
  # Add bird distribution data
  geom_point(data = Swallows_temp, aes(x = Longitude, y = Latitude,
                                       size = SpeciesTotal1,
                                       color = SpeciesTotal1,
                                       NULL)) +
  # Add some points (black) for cities
  geom_point(data = ont_cities, aes(x = long, y = lat), size = 0.8, color = "#0033cc", alpha = 0.5) +
  geom_text(data = ont_cities, aes(x = long, y = lat, label = name),hjust=1.2, vjust=0.4, size = 0.8, color = "#0033cc", alpha = 0.1) +
  # Themes
  theme_classic() +
  scale_color_gradient(guide = guide_colourbar(nbin=100, draw.ulim = FALSE, draw.llim = FALSE),
                       limits = c(0,100),
                       breaks = c(0, 20, 40, 60, 80, 100),
                       low = "#FFF15A", high = "Red",
                       name = "Birds\nRecorded",
                       labels = c(0, 20, 40, 60, 80, "100 or more")) +
  scale_size_continuous(limits = c(0, 100), # Range of number observed
                        range = c(1, 3), # Range of size of dots
                        name = "Birds Observed at Route",
                        guide = "none") +

  labs(x = "Longitude", y = "Latitude", title = Plot_Title) +
  theme(axis.text.x = element_text(size = 9, angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        # legend.key.size = unit(0.5, "cm"),
        legend.title = element_text(size = 10, vjust = 2),
        NULL) +
  NULL
```

## Mass produce .png files -- running these code chunks will slow down your computer

### Tree Swallows

```{r}
# Determine maximum of colour bar
swallows1 %>% 
  filter(species == "Tree Swallow") %>% 
  arrange(desc(SpeciesTotal)) %>% 
  head()

for (YearNo in c(1994:2015)) {
  
Swallows_temp <- swallows1 %>%
filter(Year == YearNo, species == "Tree Swallow")

# Generate plot title
Plot_Title <- paste("Ontario", unique(Swallows_temp$species), "Distribution:", unique(Swallows_temp$Year) )

# Generate plot 
plot_temp <- area_map +
  # Add bird distribution data
  geom_point(data = Swallows_temp, aes(x = Longitude, y = Latitude,
                                       size = SpeciesTotal1,
                                       color = SpeciesTotal1,
                                       NULL)) +
  # Add some points (black) for cities
  geom_point(data = ont_cities, aes(x = long, y = lat), size = 0.8, color = "#0033cc", alpha = 0.5) +
  geom_text(data = ont_cities, aes(x = long, y = lat, label = name),hjust=1.2, vjust=0.4, size = 0.8, color = "#0033cc", alpha = 0) +
  # Themes
  theme_classic() +
  scale_color_gradient(guide = guide_colourbar(nbin=100, draw.ulim = FALSE, draw.llim = FALSE),
                       limits = c(0,100),
                       breaks = c(0, 20, 40, 60, 80, 100),
                       low = "#FFF15A", high = "Red",
                       name = "Birds\nRecorded",
                       labels = c(0, 20, 40, 60, 80, "100 or more")) +
  scale_size_continuous(limits = c(0, 100), # Range of number observed
                        range = c(1, 3), # Range of size of dots
                        name = "Birds Observed at Route",
                        guide = "none") +

  labs(x = "Longitude", y = "Latitude", title = Plot_Title) +
  theme(axis.text.x = element_text(size = 9, angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        # legend.key.size = unit(0.5, "cm"),
        legend.title = element_text(size = 10, vjust = 2),
        NULL) +
  NULL

  ggsave(paste(unique(Swallows_temp$species), unique(Swallows_temp$Year), ".png"), plot = plot_temp)

}

```

### Purple Martin

```{r}
# Determine maximum of colour bar
swallows1 %>% 
  filter(species == "Purple Martin") %>% 
  arrange(desc(SpeciesTotal)) %>% 
  head()

for (YearNo in c(1994:2015)) {
  
Swallows_temp <- swallows1 %>%
filter(Year == YearNo, species == "Purple Martin")

# Generate plot title
Plot_Title <- paste("Ontario", unique(Swallows_temp$species), "Distribution:", unique(Swallows_temp$Year) )

# Generate plot 
plot_temp <- area_map +
  # Add bird distribution data
  geom_point(data = Swallows_temp, aes(x = Longitude, y = Latitude,
                                       size = SpeciesTotal1,
                                       color = SpeciesTotal1,
                                       NULL)) +
  # Add some points (black) for cities
  geom_point(data = ont_cities, aes(x = long, y = lat), size = 0.8, color = "#0033cc", alpha = 0.5) +
  geom_text(data = ont_cities, aes(x = long, y = lat, label = name),hjust=1.2, vjust=0.4, size = 0.8, color = "#0033cc", alpha = 0) +
  # Themes
  theme_classic() +
  scale_color_gradient(guide = guide_colourbar(nbin=100, draw.ulim = FALSE, draw.llim = FALSE),
                       limits = c(0,93),
                       breaks = c(0, 20, 40, 60, 80, 93),
                       low = "#FFF15A", high = "Red",
                       name = "Birds\nRecorded",
                       labels = c(0, 20, 40, 60, 80, 93)) +
  scale_size_continuous(limits = c(0, 100), # Range of number observed
                        range = c(1, 3), # Range of size of dots
                        name = "Birds Observed at Route",
                        guide = "none") +

  labs(x = "Longitude", y = "Latitude", title = Plot_Title) +
  theme(axis.text.x = element_text(size = 9, angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        # legend.key.size = unit(0.5, "cm"),
        legend.title = element_text(size = 10, vjust = 2),
        NULL) +
  NULL

  ggsave(paste(unique(Swallows_temp$species), unique(Swallows_temp$Year), ".png"), plot = plot_temp)

}

```

### Cliff Swallow

```{r}
# Determine maximum of colour bar
swallows1 %>% 
  filter(species == "Cliff Swallow") %>% 
  arrange(desc(SpeciesTotal)) %>% 
  head()

for (YearNo in c(1994:2015)) {
  
Swallows_temp <- swallows1 %>%
filter(Year == YearNo, species == "Cliff Swallow")

# Generate plot title
Plot_Title <- paste("Ontario", unique(Swallows_temp$species), "Distribution:", unique(Swallows_temp$Year) )

# Generate plot 
plot_temp <- area_map +
  # Add bird distribution data
  geom_point(data = Swallows_temp, aes(x = Longitude, y = Latitude,
                                       size = SpeciesTotal1,
                                       color = SpeciesTotal1,
                                       NULL)) +
  # Add some points (black) for cities
  geom_point(data = ont_cities, aes(x = long, y = lat), size = 0.8, color = "#0033cc", alpha = 0.5) +
  geom_text(data = ont_cities, aes(x = long, y = lat, label = name),hjust=1.2, vjust=0.4, size = 0.8, color = "#0033cc", alpha = 0) +
  # Themes
  theme_classic() +
  scale_color_gradient(guide = guide_colourbar(nbin=100, draw.ulim = FALSE, draw.llim = FALSE),
                       limits = c(0,100),
                       breaks = c(0, 20, 40, 60, 80, 100),
                       low = "#FFF15A", high = "Red",
                       name = "Birds\nRecorded",
                       labels = c(0, 20, 40, 60, 80, "100 or more")) +
  scale_size_continuous(limits = c(0, 100), # Range of number observed
                        range = c(1, 3), # Range of size of dots
                        name = "Birds Observed at Route",
                        guide = "none") +

  labs(x = "Longitude", y = "Latitude", title = Plot_Title) +
  theme(axis.text.x = element_text(size = 9, angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        # legend.key.size = unit(0.5, "cm"),
        legend.title = element_text(size = 10, vjust = 2),
        NULL) +
  NULL

  ggsave(paste(unique(Swallows_temp$species), unique(Swallows_temp$Year), ".png"), plot = plot_temp)

}

```

### Barn Swallow

```{r}
# Determine maximum of colour bar
swallows1 %>% 
  filter(species == "Barn Swallow") %>% 
  arrange(desc(SpeciesTotal)) %>% 
  head()

for (YearNo in c(1994:2015)) {
  
Swallows_temp <- swallows1 %>%
filter(Year == YearNo, species == "Barn Swallow")

# Generate plot title
Plot_Title <- paste("Ontario", unique(Swallows_temp$species), "Distribution:", unique(Swallows_temp$Year) )

# Generate plot 
plot_temp <- area_map +
  # Add bird distribution data
  geom_point(data = Swallows_temp, aes(x = Longitude, y = Latitude,
                                       size = SpeciesTotal1,
                                       color = SpeciesTotal1,
                                       NULL)) +
  # Add some points (black) for cities
  geom_point(data = ont_cities, aes(x = long, y = lat), size = 0.8, color = "#0033cc", alpha = 0.5) +
  geom_text(data = ont_cities, aes(x = long, y = lat, label = name),hjust=1.2, vjust=0.4, size = 0.8, color = "#0033cc", alpha = 0) +
  # Themes
  theme_classic() +
  scale_color_gradient(guide = guide_colourbar(nbin=100, draw.ulim = FALSE, draw.llim = FALSE),
                       limits = c(0,100),
                       breaks = c(0, 20, 40, 60, 80, 100),
                       low = "#FFF15A", high = "Red",
                       name = "Birds\nRecorded",
                       labels = c(0, 20, 40, 60, 80, "100 or more")) +
  scale_size_continuous(limits = c(0, 100), # Range of number observed
                        range = c(1, 3), # Range of size of dots
                        name = "Birds Observed at Route",
                        guide = "none") +

  labs(x = "Longitude", y = "Latitude", title = Plot_Title) +
  theme(axis.text.x = element_text(size = 9, angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        # legend.key.size = unit(0.5, "cm"),
        legend.title = element_text(size = 10, vjust = 2),
        NULL) +
  NULL

  ggsave(paste(unique(Swallows_temp$species), unique(Swallows_temp$Year), ".png"), plot = plot_temp)

}

```

### Bank Swallow

```{r}
# Determine maximum of colour bar
swallows1 %>% 
  filter(species == "Bank Swallow") %>% 
  arrange(desc(SpeciesTotal)) %>% 
  head() 
  # write_csv("temp.csv")

for (YearNo in c(1994:2015)) {
  
Swallows_temp <- swallows1 %>%
filter(Year == YearNo, species == "Bank Swallow")

# Generate plot title
Plot_Title <- paste("Ontario", unique(Swallows_temp$species), "Distribution:", unique(Swallows_temp$Year) )

# Generate plot 
plot_temp <- area_map +
  # Add bird distribution data
  geom_point(data = Swallows_temp, aes(x = Longitude, y = Latitude,
                                       size = SpeciesTotal1,
                                       color = SpeciesTotal1,
                                       NULL)) +
  # Add some points (black) for cities
  geom_point(data = ont_cities, aes(x = long, y = lat), size = 0.8, color = "#0033cc", alpha = 0.5) +
  geom_text(data = ont_cities, aes(x = long, y = lat, label = name),hjust=1.2, vjust=0.4, size = 0.8, color = "#0033cc", alpha = 0) +
  # Themes
  theme_classic() +
  scale_color_gradient(guide = guide_colourbar(nbin=100, draw.ulim = FALSE, draw.llim = FALSE),
                       limits = c(0,100),
                       breaks = c(0, 20, 40, 60, 80, 100),
                       low = "#FFF15A", high = "Red",
                       name = "Birds\nRecorded",
                       labels = c(0, 20, 40, 60, 80, "100 or more")) +
  scale_size_continuous(limits = c(0, 100), # Range of number observed
                        range = c(1, 3), # Range of size of dots
                        name = "Birds Observed at Route",
                        guide = "none") +

  labs(x = "Longitude", y = "Latitude", title = Plot_Title) +
  theme(axis.text.x = element_text(size = 9, angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        # legend.key.size = unit(0.5, "cm"),
        legend.title = element_text(size = 10, vjust = 2),
        NULL) +
  NULL

  ggsave(paste(unique(Swallows_temp$species), unique(Swallows_temp$Year), ".png"), plot = plot_temp)

}

```

### Northern Rough-winged Swallow

```{r}
swallows1 %>% 
  filter(species == "Northern Rough-winged Swallow") %>% 
  arrange(desc(SpeciesTotal)) %>% 
  head() 

for (YearNo in c(1994:2015)) {
  
Swallows_temp <- swallows1 %>%
filter(Year == YearNo, species == "Northern Rough-winged Swallow")

# Generate plot title
Plot_Title <- paste("Ontario", unique(Swallows_temp$species), "Distribution:", unique(Swallows_temp$Year) )

# Generate plot 
plot_temp <- area_map +
  # Add bird distribution data
  geom_point(data = Swallows_temp, aes(x = Longitude, y = Latitude,
                                       size = SpeciesTotal1,
                                       color = SpeciesTotal1,
                                       NULL)) +
  # Add some points (black) for cities
  geom_point(data = ont_cities, aes(x = long, y = lat), size = 0.8, color = "#0033cc", alpha = 0.5) +
  geom_text(data = ont_cities, aes(x = long, y = lat, label = name),hjust=1.2, vjust=0.4, size = 0.8, color = "#0033cc", alpha = 0) +
  # Themes
  theme_classic() +
  scale_color_gradient(guide = guide_colourbar(nbin=100, draw.ulim = FALSE, draw.llim = FALSE),
                       limits = c(0,45),
                       breaks = c(0, 10, 20, 30, 40, 45),
                       low = "#FFF15A", high = "Red",
                       name = "Birds\nRecorded",
                       labels = c(0, 10, 20, 30, 40, 45)) +
  scale_size_continuous(limits = c(0, 100), # Range of number observed
                        range = c(1, 3), # Range of size of dots
                        name = "Birds Observed at Route",
                        guide = "none") +

  labs(x = "Longitude", y = "Latitude", title = Plot_Title) +
  theme(axis.text.x = element_text(size = 9, angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        # legend.key.size = unit(0.5, "cm"),
        legend.title = element_text(size = 10, vjust = 2),
        NULL) +
  NULL

  ggsave(paste(unique(Swallows_temp$species), unique(Swallows_temp$Year), ".png"), plot = plot_temp)

}

```

### All swallows combined (plots look bad)

```{r}
swallows2 <- swallows1 %>% 
  group_by(RouteYear) %>% 
  mutate(TotalBirds = sum(SpeciesTotal))

# write_csv(swallows2, "temp.csv")

swallows2 <- swallows2 %>% 
  mutate(SpeciesTotal1 = ifelse(SpeciesTotal > 200, 200, SpeciesTotal))

for (YearNo in c(1994:2015)) {
  
Swallows_temp <- swallows2 %>%
filter(Year == YearNo)

# Generate plot title
Plot_Title <- paste("All Ontario Birds Distribution", unique(Swallows_temp$Year) )

# Generate plot 
plot_temp <- area_map +
  # Add bird distribution data
  geom_point(data = Swallows_temp, aes(x = Longitude, y = Latitude,
                                       size = SpeciesTotal1,
                                       color = SpeciesTotal1,
                                       NULL)) +
  # Add some points (black) for cities
  geom_point(data = ont_cities, aes(x = long, y = lat), size = 0.8, color = "#0033cc", alpha = 0.5) +
  geom_text(data = ont_cities, aes(x = long, y = lat, label = name),hjust=1.2, vjust=0.4, size = 0.8, color = "#0033cc", alpha = 0) +
  # Themes
  theme_classic() +
  scale_color_gradient(guide = guide_colourbar(nbin=100, draw.ulim = FALSE, draw.llim = FALSE),
                       limits = c(0,200),
                       breaks = c(0, 40, 80, 120, 160, 200),
                       low = "#FFF15A", high = "Red",
                       name = "Birds\nRecorded",
                       labels = c(0, 40, 80, 120, 160, "200 and more")) +
  scale_size_continuous(limits = c(0, 200), # Range of number observed
                        range = c(1, 3), # Range of size of dots
                        name = "Birds Observed at Route",
                        guide = "none") +

  labs(x = "Longitude", y = "Latitude", title = Plot_Title) +
  theme(axis.text.x = element_text(size = 9, angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        # legend.key.size = unit(0.5, "cm"),
        legend.title = element_text(size = 10, vjust = 2),
        NULL) +
  NULL

  ggsave(paste("All Ontario Birds Distribution", unique(Swallows_temp$Year), ".png"), plot = plot_temp)

}

```

## Animation: done using Windows Movie Maker

(End)