---
title: "R Notebook"
output: html_notebook
---
```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(hexbin)
swallows <- read_csv("Swallows Data Expanded.csv")
```


Mutating the other weather columns that consist of 'Start' and 'End' to obtain an average weather condition column for both temperature, wind, and sky conditions.
```{r}
#Mutating AvgTemp column
swallows$StartTemp <- as.numeric(swallows$StartTemp)
swallows$EndTemp <- as.numeric(swallows$EndTemp)

swallows <- swallows %>% 
  mutate(AvgTemp = ((StartTemp + EndTemp)/2))

#Mutating AvgWind column
swallows$StartWind <- as.numeric(swallows$StartWind)
swallows$EndWind <- as.numeric(swallows$EndWind)

swallows <- swallows %>% 
  mutate(AvgWind = ((StartWind + EndWind)/2))

#Mutating AvgSky column
swallows$StartSky <- as.numeric(swallows$StartSky)
swallows$EndSky <- as.numeric(swallows$EndSky)

swallows <- swallows %>% 
  mutate(AvgSky = ((StartSky + EndSky)/2))
```


Compiled plots of swallows vs both Start and End temp
```{r warning=FALSE, message=FALSE}
#This is just to get an idea of the kind of range and distribution of data we are working with 
ggplot(swallows, aes(y = SpeciesTotal))+
  geom_point(aes(x = StartTemp), colour = "orange")+
  geom_point(aes(x = EndTemp), colour = "blue", alpha = 0.25)+
  xlab("Temp")+
  ylab("Number of Swallows")
```


Compiled plots of swallows vs both Start and End Sky
```{r}
ggplot(swallows, aes(y = SpeciesTotal))+
  geom_point(aes(x = StartSky), colour = "blue4")+
  geom_point(aes(x = EndSky), colour = "cornflowerblue", alpha = 0.25)+
  xlab("Sky Conditions")+
  ylab("Number of Swallows")+
  facet_wrap(~species)
```


Compiled plots of swallows vs both Start and End Wind
```{r}
ggplot(swallows, aes(y = SpeciesTotal))+
  geom_point(aes(x = StartWind), colour = "greenyellow")+
  geom_point(aes(x = EndWind), colour = "honeydew4", alpha = 0.25)+
  xlab("Wind Conditions")+
  ylab("Number of Swallows")+
  facet_wrap(~species)
```


Out of curiosity, plotted average temperature over time to see if there are any dramatic temperature changes over the years.
```{r}
ggplot(swallows, aes(y = AvgTemp))+
  geom_point(aes(x = Year), colour = "brown2")+
  xlab("Year")+
  ylab("Average Temperature (°C)")
#It would appear the temperature doesn't vary much over the years
```


Made a linear model to see if average weather conditions are significantly affected from year to year.
```{r}
#Year-Temp Model
year_temp_model <- lm(AvgTemp ~ Year, data = swallows)
summary(year_temp_model)
```


```{r}
plot(year_temp_model)
```


```{r}
#Year-Wind Model
year_wind_model <- lm(AvgWind ~ Year, data = swallows)
summary(year_wind_model)

#So apparently average wind conditions significantly change from year to year!
```


```{r}
plot(year_wind_model)
```


```{r}
#Year-Sky Model
year_sky_model <- lm(AvgSky ~ Year, data = swallows)
summary(year_sky_model)
```


```{r}
plot(year_sky_model)
```


Plot for each Swallow species based on Latitude/Longitude. This graph has potential for making a heat map! Since you can basically almost see the outline of Southern Ontario!
```{r warning=FALSE, message=FALSE}
ggplot(swallows, aes(x = Longitude, y = Latitude))+
  geom_point(aes(colour = species))+
  facet_wrap(~species)+
  theme(axis.text.x = element_text(size = 9, angle = 45, hjust = 1))
```


Now looking at our distribution of data to see if the column mutations worked out properly.
```{r warning=FALSE, message=FALSE}
#TEMP
qplot(AvgTemp, SpeciesTotal, data = swallows)+
  xlab("Average Temperature (°C)")+
  ylab("Number of Swallows")
```


```{r}
#WIND
qplot(AvgWind, SpeciesTotal, data = swallows)+
  xlab("Average Wind")+
  ylab("Number of Swallows")
```


```{r}
#SKY
qplot(AvgSky, SpeciesTotal, data = swallows)+
  xlab("Average Sky")+
  ylab("Number of Swallows")
```


Linear models for Swallows and AvgTemp/Wind
```{r}
#TEMP
swallows_temp_model <- lm(SpeciesTotal ~ AvgTemp, data = swallows)
summary(swallows_temp_model)
```


```{r}
plot(swallows_temp_model)
```


```{r}
#WIND
swallows_wind_model <- lm(SpeciesTotal ~ AvgWind, data = swallows)
summary(swallows_wind_model)

#This is backed by our previous linear model that shows there are significant changes in wind from year to year. 
```


```{r}
plot(swallows_wind_model)
```


#ANOVA
To test the linear models of temp and wind
```{r}
#TEMP
anova(swallows_temp_model)
```


```{r}
#WIND
anova(swallows_wind_model)
```


Sky isn't necessarily a super important variable that we want to look at but I will do a linear model and ANOVA for it out of curiosity to see how it impacts swallows. 
Linear model for sky conditions
```{r}
swallows_sky_model <- lm(SpeciesTotal ~ AvgSky, data = swallows)
summary(swallows_sky_model)
```


ANOVA of sky conditions
```{r}
anova(swallows_sky_model)
#So apparently sky conditions have some significance to them  -- good thing I looked at it
```


Hexbin Plot of Swallows vs Temp, Wind, and Sky
```{r warning=FALSE, message=FALSE}
#TEMP
ggplot(swallows, aes(AvgTemp, SpeciesTotal, na.rm = T))+
  geom_hex()+
  xlab("Average Temperature (°C)")+
  ylab("Number of Swallows")
```


```{r}
#WIND
ggplot(swallows, aes(AvgWind, SpeciesTotal, na.rm = T))+
  geom_hex()+
  xlab("Average Wind")+
  ylab("Number of Swallows")
```


```{r}
#SKY
ggplot(swallows, aes(AvgSky, SpeciesTotal, na.rm = T))+
  geom_hex()+
  xlab("Average Sky")+
  ylab("Number of Swallows")
```


Plots showing the normal distribution of the swallows depending on the average weather conditions
```{r warning=FALSE, message=FALSE}
ggplot(swallows, aes(x = swallows$AvgTemp)) + 
	geom_histogram(aes(y = ..density..), binwidth = 0.5) + 
	stat_function(fun=dnorm,
	              color="red",
	              lwd = 1,
	              args=list(mean = mean(swallows$AvgTemp, na.rm = T),
	                        sd = sd(swallows$AvgTemp, na.rm = T)))+
  xlab("Average Temperature (°C)")+
  ylab("Frequency")
```


```{r warning=FALSE, message=FALSE}
ggplot(swallows, aes(x = swallows$AvgSky)) + 
	geom_histogram(aes(y = ..density..), binwidth = 0.5) + 
	stat_function(fun=dnorm,
	              color="red",
	              lwd = 1,
	              args=list(mean = mean(swallows$AvgSky, na.rm = T),
	                        sd = sd(swallows$AvgSky, na.rm = T)))+
  xlab("Average Sky")+
  ylab("Frequency")
```


```{r warning=FALSE, message=FALSE}
ggplot(swallows, aes(x = swallows$AvgWind)) + 
	geom_histogram(aes(y = ..density..), binwidth = 0.5) + 
	stat_function(fun=dnorm,
	              color="red",
	              lwd = 1,
	              args=list(mean = mean(swallows$AvgWind, na.rm = T),
	                        sd = sd(swallows$AvgWind, na.rm = T)))+
  xlab("Average Wind")+
  ylab("Frequency")
```


```{r}
swallows_temp_wind_sky_model <- lm(SpeciesTotal ~ AvgTemp + AvgWind + AvgSky, data = swallows)
summary(swallows_temp_wind_sky_model)
```


```{r}
anova(swallows_temp_wind_sky_model)
```


```{r}
plot(swallows_temp_wind_sky_model)
```


Now to visualize and analyze the impact cars have on swallow abundance. Car_Total will be our proxy for human disturbance 
```{r warning=FALSE, message=FALSE}
#For some reason it treats Car_Total as a factor
swallows$Car_Total <- as.numeric(swallows$Car_Total)

swallows %>% 
  filter(!is.na(SpeciesTotal)) %>% 
  ggplot(., aes(x = Car_Total, y = SpeciesTotal))+
  geom_point(alpha = 0.5)+
  coord_cartesian(ylim = c(0, 500), xlim = c(0,500))+
  ylab("Number of Swallows")+
  xlab("Number of Cars")+
  facet_wrap(~species)

#So it looks like cars might have something to do with swallow numbers since more cars = less swallows
```


```{r}
swallows_cars_model <- lm(SpeciesTotal ~ Car_Total, data = swallows)
summary(swallows_cars_model)
#Apparently traffic does not seem to affect swallows!
```


```{r}
plot(swallows_cars_model)
```


Run an ANOVA just to see if the linear model's findings are supported
```{r}
anova(swallows_cars_model)
#Cars/human disturbance really don't significantly impact swallows
```

